Eco-evolutionary dynamics of finite populations from first principles

### 2023-03-07 ind_files/04_unification_discussion.tex
Add some references for SPDE related material

where $N_K = \sum x_i$ is the total population size scaled by $K$ (and thus $KN_K$ is the total population size), $A_i^{+} = x_i\tau_i(\mathbf{x}) + \mu Q_i(\mathbf{x})$, and each $B^{(i)}_t$ is an independent one-dimensional standard Brownian motion. Equation \eqref{nD_stochastic_RM} is in `replicator-mutator' form, and letting $K \to \infty$ recovers the standard replicator-mutator equation. The first term represents the direct effects of forces captured in classic deterministic models, and reflects a selection-mutation balance. However, finite populations experience a new directional force dependent on $\tau_i(\mathbf{x})$, the per-capita turnover rate of type $i$, that cannot be captured in infinite population models \citep{kuosmanen_turnover_2022}. Remarkably, this term acts in a way that is mathematically identical to the classical action of selection and mutation in infinite population models as captured by the first term in \eqref{nD_stochastic_RM}, but in the opposite direction - A higher relative $\tau_i$ leads to a decrease in frequency (Notice the minus sign before the second term in \eqref{nD_stochastic_RM}).

@@

\eqref{nD_stochastic_RM} decompose into a selection term $\textrm{Cov}(\tau,f)$ for reduced turnover rates and a transmission bias term that vanishes in the weak mutation ($\mu \to 0$) limit. Following \cite{week_white_2021}, we refer to the effect of the covariance (the second term of equation \eqref{nD_stochastic_Price}) as \emph{noise-induced selection} since it occurs exactly analogously to classical natural selection (but for lower $\tau$) and is induced purely by the finiteness of the population. Since this evolutionary force is unique to finite populations and has therefore been overlooked in classical population genetics, it warrants some more detailed discussion. Biologically, the $\textrm{Cov}(\tau,f)$ term (with a negative sign) describes a biasing effect due to differential turnover rates and can intuitively be understood as being similar to gambler's ruin in probability theory through the following reasoning: If a type $i$ has a higher $\tau_i$, it experiences greater turnover due to a generally higher birth and death rate and thus experience more births and deaths in a given time interval than an otherwise equivalent species with a lower $\tau_i$. More events mean greater demographic stochasticity, and types with a higher $\tau_i$ thus tend to be eliminated by a stochastic analog of selection because they experience more chance events (births and deaths) in a given time period. This effect is less visible if the total population size is higher because larger populations generally experience less stochasticity, which is reflected in the $1/N_K$ factor in this term. This stochastic analog of selection for reduced turnover rates, captured by the second term of equation \eqref{nD_stochastic_RM}, is the force responsible for the `reversal of the direction of deterministic selection' induced by demographic noise in previous studies \citep{houchmandzadeh_selection_2012, houchmandzadeh_fluctuation_2015, constable_demographic_2016, mcleod_social_2019}. Note that types that tend to increase the \emph{total} population size $KN_K(t)$ (such as altruists in evolutionary theory and mutualists in ecological communities) will reduce the magnitude of this effect compared to types that do not facilitate such an increase, such as cheaters and highly competitive species, which could explain why this effect preferentially favors the former types in reversing the direction of deterministic selection in finite population models with fluctuating population sizes \citep{houchmandzadeh_fluctuation_2015, chotibut_evolutionary_2015, mcleod_social_2019}. The fact that total population size controls the strength of noise-induced selection also explains why cooperation is favoured in the early transient period of population growth \citep{melbinger_evolutionary_2010} when simulations are initiated from a small population size - In the early transient period, $N_K(t)$ is small, and the biasing effect of differential turnover rates is stronger, thus favouring cooperation. More generally, selection for reduced turnover rate could explain why cooperation often persists for a long time in finite population IbMs (and the real world) despite infinite population models predicting their extinction. The fact that the entire term is multiplied by $(KN_K(t))^{-1}$ suggests that the effect of this force is weak for medium to large populations, which explains why the persistence of cooperators is often only observed in restrictive sounding conditions such as quasi-neutrality, rapid attraction to a slow manifold, or a weak selection + weak mutation limit \citep{mcleod_social_2019}. In all three of these cases, the first term on the RHS of \eqref{nD_eqn_for_frequencies} becomes identically 0. It therefore no longer contributes to the trait frequency dynamics, thus allowing us to see the contributions of the second term.\\

@@

Note that just like the statistical mean, the statistical variance is a random variable and is not to be confused with the probabilistic/ensemble variance. Upon substituting $f = w$ in \eqref{nD_stochastic_Price} and taking expectations over the underlying probability space, we obtain:

@@

Taking $K \to \infty$ in \eqref{nD_stochastic_Fisher} recovers a well-known equation in population genetics upon noting that the process tends to a deterministic process as $K \to \infty$, as noted in section \ref{sec_nD_det_limit}, and thus the expectation value in the infinite population case is superfluous.The first term, $\sigma^2_w$, is the subject of Fisher's fundamental theorem \citep{fisher_genetical_1930,  price_fishers_1972, frank_fishers_1992, kokko_stagnation_2021}. The second term of equation \eqref{nD_stochastic_Fisher} is a manifestation of noise-induced selection and vanishes in the infinite population limit, and is thus particular to finite populations. Finally, the last term arises in both finite and infinite populations whenever $w_i$ can vary over time \citep{baez_fundamental_2021}, be it through frequency-dependent selection, phenotypic plasticity, varying environments, or other ecological mechanisms, and represents feedback effects on the fitness $w_i$ of the $i\textsuperscript{th}$ species over short (`ecological') time-scales. The fact that Fisher appears to have been rather vague and dismissive of this feedback \citep{fisher_genetical_1930} has led to much discussion, debate, and confusion about the interpretation, importance, and implications of his `fundamental theorem' (see \cite{kokko_stagnation_2021} and sources cited therein).\\

@@

where $w(x|\phi)$ and $\tau(x|\phi)$ are functionals that respectively the Malthusian fitness and per-capita turnover rate of type $x \in \mathcal{T}$ (now a continuous variable) in a population $\phi$. We then showed that this equation yields some well-known frameworks of quantitative genetics in the infinite population ($K \to \infty$) limit, thus illustrating consistency with known theories. If the `intuitive' version of It\^{o}'s formula for $\mathcal{M}_K(\mathcal{T})$ valued stochastic processes obtained by `taking the limit' $m \to \infty$ in the It\^{o}'s formula for $m$-dimensional stochastic processes is true, then the exact same steps carried out in Appendices \ref{App_density_to_freq} and \ref{App_stoch_var_eqns} will `go through' essentially unchanged for quantitative traits and yield the equations obtained by simply taking $m \to \infty$ in \eqref{nD_stochastic_RM}, \eqref{nD_stochastic_Price} and \eqref{nD_stochastic_Price_variance} (with sums replaced by integrals and derivatives replaced by functional derivatives as appropriate) as the `fundamental theorems' for the evolution of quantitative traits. Indeed, \cite{week_white_2021} have recently proposed exactly the It\^o formula we would need via a heuristic It\^o multiplication table for $L^2(\mathbb{R}, m)$ valued stochastic processes with a.s. finite Hilbert-Schmidt norm and have shown that their heuristics are equivalent to the rigorous infinite-dimensional stochastic calculus proposed by \cite{da_prato_stochastic_2014} for more general Hilbert space valued processes\footnote{If this sentence sounds like arcane gibberish to you, don't worry too much about the details - The essence is that \cite{week_white_2021} have proven the `correct' formula we need holds whenever $b(x|\nu), d(x|\nu)$ and $\mathcal{M}_K(\mathcal{T})$ together fulfill certain technical requirements. My heuristics here are only a `first step' and have focused on accessibility over mathematical propriety, and thus make no attempt to verify whether and when these technical requirements are satisfied.}. Furthermore, using these `spacetime white noise heuristics' together with a particular functional form for (weak) solutions to certain SPDEs, \cite{week_white_2021} have arrived at exactly the infinite-dimensional version of equations \eqref{nD_stochastic_Price} and \eqref{nD_stochastic_Price_variance} from the `SPDE side'\footnote{If the well-known equivalence between SDEs and Fokker-Planck equations (see section \ref{sec_math_background}) carries through to infinite-dimensions, it becomes an equivalence between SPDEs and functional Fokker-Planck equations.} under certain additional assumptions required to ensure existence/uniqueness of solutions to the relevant SPDEs. Equation \eqref{disc_functional_field_eqns} can thus be argued to be the (functional) Fokker-Planck view of the stochastic processes for the evolutionary ecology of quantitative traits recently studied in \cite{week_white_2021}. Like \cite{week_white_2021}, I believe that rigorously establishing the precise criteria needed for general existence and uniqueness of solutions to equations of the form \eqref{disc_functional_field_eqns} (or the SPDE versions presented in \cite{week_white_2021}, which resemble our equation \eqref{functional_langevin}) presents an important mathematical direction for future work that may yield new insights into both the biology of quantitative traits and the mathematics of measure-valued stochastic processes.\\
More generally, the intimate relation between It\^o SDEs and Fokker-Planck equations has been extensively exploited in the applied mathematics literature \citep{van_kampen_stochastic_1981,oksendal_stochastic_1998, gardiner_stochastic_2009} since different problems are more appropriately attacked from one side than the other. I too have repeatedly used this equivalence in chapter \ref{chap_BD} to arrive at our final equations. In the infinite-dimensional case, the functional Fokker-Planck equation \eqref{disc_functional_field_eqns} (the `physics' side) allows us to attack questions of phenotypic clustering and evolutionary branching via powerful tools such as Fourier analysis (\cite{rogers_demographic_2012, rogers_modes_2015}; See also Appendix \ref{App_examples} for a general model-independent pedagogical derivation) that would have been significantly harder if one only used SPDEs. On the other hand, some questions are much more easily handled using tools from the `math' side of SPDEs such as infinitesimal generators and the martingale problem (see \cite{champagnat_unifying_2006} or \cite{week_white_2021}) rather than being forced to work with functional Fokker-Planck equations. Rigorously establishing a relation between functional Fokker-Planck equations of the form \eqref{disc_functional_field_eqns} and SPDEs of the form \eqref{functional_langevin} (or those studied in \cite{week_white_2021}) for measure-valued branching processes could thus prove very fruitful for developing more integrative eco-evolutionary theory since it allows us to seamlessly transition between two alternative views of the same object. Importantly, the formalism I develop here in terms of functional Fokker-Planck equations likely does \emph{not} carry over to the study of higher dimensional quantitative traits (or populations which vary in two or more one-dimensional quantitative traits) because the probability density function $P(\phi,t)$ likely does not exist in two or more dimensions \citep{fleming_measure-valued_1979,champagnat_unifying_2006}. My understanding is that the evolution of quantitative traits is also very difficult to study in two or more dimensions from the SPDE side for largely the same reason, and overcoming this difficulty may need new approaches altogether.\\
To the best of my knowledge, a general formulation ot stochastic field equations from the functional Fokker-Planck side in the manner I have carried out here has not been done before. Similar equations have been formulated for some specific stochastic models of quantitative trait evolution \citep{rogers_demographic_2012,rogers_modes_2015} and population ecology of size-structured communities \citep{odwyer_integrative_2009}. Stochastic field equations are also known in mathematical neurobiology \citep{buice_field-theoretic_2007,bressloff_stochastic_2010,coombes_neural_2014}, and have recently been proposed in a model of collective motion \citep{o_laighleis_minimal_2018}, suggesting that the study and formalization of such equations by pure mathematicians could have wide-spread applications. As is, such field equations are not on very solid mathematical footing, and are currently primarily used by physicists working in areas such as statistical field theory and are largely attacked using ingenious (and often heuristic) arguments and tools in a somewhat system-specific manner. Equation \eqref{disc_functional_field_eqns} opens up the study of quantitative trait dynamics in finite fluctuating populations to analysis using various powerful tools of physics and stochastic processes such as duality \citep{greenman_duality_2020}, the Fock space representation \citep{dodd_many-body_2009, del_razo_probabilistic_2022}, and the path integral formalism \citep{doi_second_1976, peliti_path_1985, dodd_many-body_2009, chow_path_2015, weber_master_2017}. Equation \eqref{disc_functional_field_eqns} also provides a more accessible alternative formulation that does not (explicitly) rely on tools from relatively `advanced' mathematical fields such as martingale theory and measure-valued processes that are typically used to study the evolution of quantitative traits in finite, fluctuating populations \citep{dawson_stochastic_1975,fleming_measure-valued_1979,champagnat_unifying_2006,etheridge_mathematical_2011,champagnat_polymorphic_2011,week_white_2021}, thus hopefully making the field more accessible to theorists without formal training in mathematical areas such as measure theory and stochastic processes.

>>>>

where $N_K = \sum x_i$ is the total population size scaled by $K$ (and thus $KN_K$ is the total population size), $A_i^{+} = x_i\tau_i(\mathbf{x}) + \mu Q_i(\mathbf{x})$, and each $B^{(i)}_t$ is an independent one-dimensional standard Brownian motion. Equation \eqref{nD_stochastic_RM} is in `replicator-mutator' form, and letting $K \to \infty$ recovers the standard replicator-mutator equation in the infinite population limit. The first term represents the direct effects of forces captured in classic deterministic models, and reflects a selection-mutation balance. However, finite populations experience a new directional force dependent on $\tau_i(\mathbf{x})$, the per-capita turnover rate of type $i$, that cannot be captured in infinite population models \citep{kuosmanen_turnover_2022}. Remarkably, this term acts in a way that is mathematically identical to the classical action of selection and mutation in infinite population models as captured by the first term in \eqref{nD_stochastic_RM}, but in the opposite direction - A higher relative $\tau_i$ leads to a decrease in frequency (Notice the minus sign before the second term in \eqref{nD_stochastic_RM}).

@@

\eqref{nD_stochastic_RM} decompose into a selection term $\textrm{Cov}(\tau,f)$ for reduced turnover rates and a transmission bias term that vanishes in the weak mutation ($\mu \to 0$) limit. Following \cite{week_white_2021}, we refer to the effect of the covariance (the second term of equation \eqref{nD_stochastic_Price}) as \emph{noise-induced selection} since it occurs exactly analogously to classical natural selection (but for lower $\tau$) and is induced purely by the finiteness of the population. Since this evolutionary force is unique to finite populations and has therefore been overlooked in classical population genetics, it warrants some more detailed discussion. Biologically, the $\textrm{Cov}(\tau,f)$ term (with a negative sign) describes a biasing effect due to differential turnover rates and can intuitively be understood as being similar to gambler's ruin in probability theory through the following reasoning: If a type $i$ has a higher $\tau_i$, it experiences greater turnover due to a generally higher birth and death rate and thus experience more births and deaths in a given time interval than an otherwise equivalent species with a lower $\tau_i$. More events mean greater demographic stochasticity, and types with a higher $\tau_i$ thus tend to be eliminated by a stochastic analog of selection because they experience more chance events (births and deaths) in a given time period. This effect is less visible if the total population size is higher because larger populations generally experience less stochasticity, which is reflected in the $1/N_K$ factor in this term. This stochastic analog of selection for reduced turnover rates, captured by the second term of equation \eqref{nD_stochastic_RM}, is the force responsible for the `reversal of the direction of deterministic selection' induced by demographic noise in previous studies \citep{houchmandzadeh_selection_2012, houchmandzadeh_fluctuation_2015, constable_demographic_2016, mcleod_social_2019}. Note that types that tend to increase the \emph{total} population size $KN_K(t)$ (such as altruists in evolutionary theory and mutualists in ecological communities) will reduce the magnitude of this effect compared to types that do not facilitate such an increase, such as cheaters and highly competitive species, which could explain why this effect preferentially favors the former types in reversing the direction of deterministic selection in finite population models with fluctuating population sizes \citep{houchmandzadeh_fluctuation_2015, chotibut_evolutionary_2015, mcleod_social_2019}. The fact that total population size controls the strength of noise-induced selection also explains why cooperation is favoured in the early transient period of population growth \citep{melbinger_evolutionary_2010} when simulations are initiated from a small population size - In the early transient period, $N_K(t)$ is small, and the biasing effect of differential turnover rates is stronger, thus favouring cooperation. More generally, selection for reduced turnover rate could explain why cooperation often persists for a long time in finite population IbMs (and the real world) despite infinite population models predicting their extinction. The fact that the entire term is multiplied by $(KN_K(t))^{-1}$ suggests that the effect of this force is weak for medium to large populations, which explains why the persistence of cooperators is often only observed under some specific conditions such as quasi-neutrality, rapid attraction to a slow manifold (\emph{i.e.} separation of ecological and evolutionary timescales), or a weak selection + weak mutation limit \citep{mcleod_social_2019}. In all three of these cases, the first term on the RHS of \eqref{nD_eqn_for_frequencies} becomes identically 0. It therefore no longer contributes to the trait frequency dynamics, thus allowing us to see the (otherwise weak) contributions of the second term.\\

@@

It is important to remember once again that just like the statistical mean, the statistical variance $\sigma^2_{f}(t)$ of a type-level quantity $f$ is a random variable obtained by calculating the variance of the quantity \emph{in the population} at time $t$, and is not to be confused with the probabilistic/ensemble variance obtained by calculating the variance of a quantity \emph{over different realizations} of the stochastic process (see section \ref{sec_stat_measures}). Upon substituting $f = w$ in \eqref{nD_stochastic_Price} and taking expectations over the underlying probability space, we obtain:

@@

Taking $K \to \infty$ in \eqref{nD_stochastic_Fisher} recovers a well-known equation in population genetics upon noting that the process tends to a deterministic process as $K \to \infty$, as noted in section \ref{sec_nD_det_limit}, and thus the expectation value in the infinite population case is superfluous. The first term, $\sigma^2_w$, is the subject of Fisher's fundamental theorem \citep{fisher_genetical_1930,  price_fishers_1972, frank_fishers_1992, kokko_stagnation_2021}. The second term of equation \eqref{nD_stochastic_Fisher} is a manifestation of noise-induced selection and vanishes in the infinite population limit, and is thus particular to finite populations. Finally, the last term arises in both finite and infinite populations whenever $w_i$ can vary over time \citep{frank_fishers_1992,kokko_stagnation_2021,baez_fundamental_2021}, be it through frequency-dependent selection, phenotypic plasticity, varying environments, or other ecological mechanisms, and represents feedback effects on the fitness $w_i$ of the $i\textsuperscript{th}$ species over short (`ecological') time-scales. The fact that Fisher appears to have been rather vague and dismissive of this feedback \citep{fisher_genetical_1930} has led to much discussion, debate, and confusion about the interpretation, importance, and implications of his `fundamental theorem' (see \cite{kokko_stagnation_2021} and sources cited therein).\\

@@

where $w(x|\phi)$ and $\tau(x|\phi)$ are functionals that respectively the Malthusian fitness and per-capita turnover rate of type $x \in \mathcal{T}$ (now a continuous variable) in a population $\phi$. We then showed that this equation yields some well-known frameworks of quantitative genetics in the infinite population ($K \to \infty$) limit, thus illustrating consistency with known theories. If the `intuitive' version of It\^{o}'s formula for $\mathcal{M}_K(\mathcal{T})$ valued stochastic processes obtained by `taking the limit' $m \to \infty$ in the It\^{o}'s formula for $m$-dimensional stochastic processes is true, then the exact same steps carried out in Appendices \ref{App_density_to_freq} and \ref{App_stoch_var_eqns} will `go through' essentially unchanged for quantitative traits and yield the equations obtained by simply taking $m \to \infty$ in \eqref{nD_stochastic_RM}, \eqref{nD_stochastic_Price} and \eqref{nD_stochastic_Price_variance} (with sums replaced by integrals and derivatives replaced by functional derivatives as appropriate) as the `fundamental theorems' for the evolution of quantitative traits. Indeed, \cite{week_white_2021} have recently proposed exactly the It\^o formula we would need via a heuristic It\^o multiplication table for $L^2(\mathbb{R}, m)$ valued stochastic processes with a.s. finite Hilbert-Schmidt norm and have shown that their heuristics are equivalent to the rigorous infinite-dimensional stochastic calculus proposed by \cite{da_prato_stochastic_2014} for more general Hilbert space valued processes\footnote{If this sentence sounds like arcane gibberish to you, don't worry too much about the details - The essence is that \cite{week_white_2021} have proven the `correct' formula we need holds whenever $b(x|\nu), d(x|\nu)$ and $\mathcal{M}_K(\mathcal{T})$ together fulfill certain technical requirements. My heuristics here are only a `first step' and have focused on accessibility over mathematical propriety, and thus make no attempt to verify whether and when these technical requirements are satisfied.}. Furthermore, using these `spacetime white noise heuristics' together with a particular functional form for (weak) solutions to certain SPDEs, \cite{week_white_2021} have arrived at precisely the infinite-dimensional version of equations \eqref{nD_stochastic_Price} and \eqref{nD_stochastic_Price_variance} obtained by `taking the limit' $m \to \infty$ in \eqref{nD_stochastic_Price} and \eqref{nD_stochastic_Price_variance} from the `SPDE side'\footnote{If the well-known equivalence between SDEs and Fokker-Planck equations (see section \ref{sec_math_background}) carries through to infinite-dimensions, it becomes an equivalence between SPDEs and functional Fokker-Planck equations.} under certain additional assumptions required to ensure existence/uniqueness of solutions to the relevant SPDEs. Equation \eqref{disc_functional_field_eqns} can thus be argued to be the (functional) Fokker-Planck view of the stochastic processes for the evolutionary ecology of quantitative traits recently studied in \cite{week_white_2021}. Like \cite{week_white_2021}, I believe that rigorously establishing the precise criteria needed for general existence and uniqueness of solutions to equations of the form \eqref{disc_functional_field_eqns} (or the SPDE versions presented in \cite{week_white_2021}, which resemble our equation \eqref{functional_langevin}) presents an important mathematical direction for future work that may yield new insights into both the biology of quantitative traits and the mathematics of measure-valued stochastic processes.\\
More generally, the intimate relation between It\^o SDEs and Fokker-Planck equations has been extensively exploited in the applied mathematics literature \citep{van_kampen_stochastic_1981,oksendal_stochastic_1998, gardiner_stochastic_2009} since different problems are more appropriately attacked from one side than the other. I too have repeatedly used this equivalence in chapter \ref{chap_BD} to arrive at our final equations. In the infinite-dimensional case, the functional Fokker-Planck equation \eqref{disc_functional_field_eqns} (the `physics' side) allows us to attack questions of phenotypic clustering and evolutionary branching via powerful tools such as Fourier analysis (\cite{rogers_demographic_2012, rogers_modes_2015}; See also Appendix \ref{App_examples} for a general model-independent pedagogical derivation) that would have been significantly harder if one only used SPDEs. On the other hand, some questions are much more easily handled using tools from the `math' side of SPDEs such as infinitesimal generators and the martingale problem (see \cite{champagnat_unifying_2006} or \cite{week_white_2021}) rather than being forced to work with functional Fokker-Planck equations. Rigorously establishing a relation between functional Fokker-Planck equations of the form \eqref{disc_functional_field_eqns} and SPDEs of the form \eqref{functional_langevin} (or those studied in \cite{week_white_2021}) for measure-valued branching processes could thus prove very fruitful for developing more integrative eco-evolutionary theory since it allows us to seamlessly transition between two alternative views of the same object. Importantly, the formalism I develop here in terms of functional Fokker-Planck equations likely does \emph{not} carry over to the study of higher dimensional quantitative traits (or populations which vary in two or more one-dimensional quantitative traits) because these processes are routinely badly behaved in higher dimensions: In particular, a probability density function $P(\phi,t)$ frequently does not exist in higher dimensions if one has any biologically non-trivial features such as interactions between types (\cite{fleming_measure-valued_1979, walsh_introduction_1986}; Also see for example \cite{evans_measure-valued_1994}). My (admittedly limited) understanding is that such processes are also rather difficult to study in two or more dimensions from the SPDE side for similar reasons --- one is forced to work with distribution valued (rather than function valued) solutions, calling for considerable amount of advanced functional analysis to make progress \citep{walsh_introduction_1986,carmona_stochastic_1999,balan_gentle_2018}.\\
To the best of my knowledge, a general formulation of stochastic field equations for the population dynamics of quantitative traits from the functional Fokker-Planck side in the manner I have carried out here has not been done before. Similar equations have been formulated for some specific stochastic models of quantitative trait evolution \citep{rogers_demographic_2012,rogers_modes_2015} and population ecology of size-structured communities \citep{odwyer_integrative_2009}. Stochastic field equations are also known in mathematical neurobiology \citep{buice_field-theoretic_2007,bressloff_stochastic_2010,coombes_neural_2014}, and have recently been proposed in a model of collective motion \citep{o_laighleis_minimal_2018}, suggesting that the study and formalization of such equations by pure mathematicians could have wide-spread applications. As is, such field equations are not on very solid mathematical footing, and are currently primarily used by physicists working in areas such as statistical field theory and are largely attacked using ingenious (and often heuristic) arguments and tools in a somewhat system-specific manner. Equation \eqref{disc_functional_field_eqns} opens up the study of quantitative trait dynamics in finite fluctuating populations to analysis using various powerful tools of physics and stochastic processes such as duality \citep{greenman_duality_2020}, the Fock space representation \citep{dodd_many-body_2009, del_razo_probabilistic_2022}, and the path integral formalism \citep{doi_second_1976, peliti_path_1985, dodd_many-body_2009, chow_path_2015, weber_master_2017}. Equation \eqref{disc_functional_field_eqns} also provides a more accessible alternative formulation that does not (explicitly) rely on tools from relatively `advanced' mathematical fields such as martingale theory and measure-valued processes that are typically used to study the evolution of quantitative traits in finite, fluctuating populations \citep{dawson_stochastic_1975,fleming_measure-valued_1979,ethier_markov_1986,champagnat_unifying_2006,etheridge_mathematical_2011,week_white_2021}, thus hopefully making the field more accessible to theorists without formal training in mathematical areas such as measure theory and stochastic processes.
